---
title: "Week 6 Homework"
author: "Jake Reichard"
format: html
toc: true
toc-depth: 4
---

# **Introduction**

Sampling was conducted in Maunalua Bay on the island of Oahu. Sensors were towed via a kayak to collect various environmental data.

**Parameters collected**

* Temperature
* Salinity
* Depth

### **Load Libraries**

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(here)
library(lubridate)
library(viridis)
library(kableExtra)
```

### **Load in data**

*Remember to check the first rows of data*

```{r}
#| output: false

conddat <- read.csv(here("week_05","data","CondData.csv"))
depthdat <- read.csv(here("week_05", "data","DepthData.csv"))

head(conddat)
head(depthdat)
```

### **Format the data** 
```{r}
#format the date to be ymd_hs
conddat$date <- mdy_hms(conddat$date) 
depthdat$date <- ymd_hms(depthdat$date)

#round cond data to the nearest 10 seconds
conddat <- conddat %>% 
  mutate(rounded_date = round_date(date, "10 second"))

#join data that only has matching data so there is no na
joined_dat <- conddat %>% 
  inner_join(depthdat, by = c("rounded_date" ="date")) #joining with the rounded date column
```

### **Data Summarization**

#### Mean by Minute
```{r}
#mean of depth temperature and salinity by min
dat_summary_min <- joined_dat %>%
  mutate( minute =floor_date(date, "minute")) %>% 
  group_by(minute) %>% 
  summarise( temp_mean = mean(Temperature, na.rm =T),
             sal_mean = mean(Salinity, na.rm =T),
             depth_mean = mean(Depth, na.rm =T),
             .groups = "drop")
```

#### Mean by Day
```{r}
#| label: tbl-mean
#| tbl-cap: "The mean temperature, salinity, and depth for the day."
#mean by day
joined_dat %>%
  mutate(day = floor_date(date, "day")) %>% 
  group_by(day) %>% 
  summarise("Mean Temperature (C)" = mean(Temperature, na.rm = T),
            "Mean Salinity" = mean(Salinity, na.rm = T),
            "Mean Depth (m)" = mean(Depth, na.rm = T),
            .groups = "drop") %>% 
  kbl() %>%  #make it a kable table
  kable_classic() %>% #adds a theme to table
  row_spec(1, bold = TRUE, color = "black", background = "lightblue") %>%  #how to highlight a row and add color
  kable_styling(full_width = F) %>%  #doesn't make it too wide
column_spec(1:4, border_right = TRUE) #adds border to the right
```


@tbl-mean has information on the mean temperature, salinity, and depth that were observed for the day.

#### Mean ploted over time
```{r}
#| warning: false
#| fig-width: 7
#| fig-height: 8
#| fig-align: "center"
#| label: fig-mean
#| fig-cap: "This is a figure shows the mean parameters over time."

#pivot to longer to graph multiple variable across time
dat_sum_long <- dat_summary_min %>% 
  pivot_longer(cols = temp_mean:depth_mean,
               names_to = "Variables",
               values_to = "Values")

#make cleaner labels for figure, will insert it into facet wrap 
variable_labels <- c(temp_mean = "Temperature (°C)",
  sal_mean = "Salinity (PSU)",
  depth_mean = "Depth (m)")

#make a plot that plots temp, sal, and depth over time
print(  
 ggplot(dat_sum_long, aes(x = minute, y = Values, color = Variables)) +
  geom_line() +
  geom_point(alpha = 0.7)+
  facet_wrap(~ Variables, scales ="free_y", ncol = 1, #makes each scale to each variables y values
             labeller = labeller(Variables = variable_labels)) +
  
  #make labels and titles 
  labs(title = "Temperature, Salinity, and Depth Over Time",
         subtitle = "Variables are averaged at one minute intervals",
         caption =  "Data sources:Location -Maunalua Bay, Silbiger et al. 2020 ",
       x = "Time", 
       y = "") +
  theme_bw()+
  theme(legend.position = "none")
)
 
```

When you look at @fig-mean we can see that when depth is relatively constant so are salinity and temperature.